---
title: "Data Manipulation in R with dplyr"
author: "廖昱嘉"
date: "2018年7月23日"
output: html_document
---
# Data Manipulation in R with dplyr   
```{r}
library(dplyr)
library(hflights)
head(hflights)
lut <- c("AA" = "American", "AS" = "Alaska", "B6" = "JetBlue", "CO" = "Continental", 
         "DL" = "Delta", "OO" = "SkyWest", "UA" = "United", "US" = "US_Airways", 
         "WN" = "Southwest", "EV" = "Atlantic_Southeast", "F9" = "Frontier", 
         "FL" = "AirTran", "MQ" = "American_Eagle", "XE" = "ExpressJet", "YV" = "Mesa")

# Add the Carrier column to hflights
hflights$Carrier <- lut[hflights$UniqueCarrier]
```   
A tbl (pronounced tibble) is just a special kind of data.frame. They make your data easier to look at, but also easier to work with.   
以下介紹五個重要的dplyr函數   
select(), which returns a subset of the columns,    
select(df, var1, var2)    
You can also use : to select a range of variables and - to exclude some variables, similar to indexing a data.frame with square brackets.    
select(df, 1:4, -2)   
select() does not change the data frame it is called on; you have to explicitly assign the result of select() to a variable to store the result.    
```{r}
head(select(hflights,ActualElapsedTime, AirTime, ArrDelay , DepDelay))
head(select(hflights,Origin:Cancelled))
head(select(hflights,1:21,-(DepTime:AirTime)))
```
Helper functions for variable selection   
starts_with("X"): every name that starts with "X",    
ends_with("X"): every name that ends with "X",    
contains("X"): every name that contains "X",    
matches("X"): every name that matches "X", where "X" can be a regular expression,   
num_range("x", 1:5): the variables named x01, x02, x03, x04 and x05,    
one_of(x): every name that appears in x, which should be a character vector.   
```{r}
head(select(hflights, ends_with("Delay")))
head(select(hflights, UniqueCarrier, ends_with("Num"), starts_with("Cancel")))
head(select(hflights, contains("Tim"), contains("Del")))
```
filter(), that is able to return a subset of the rows,    
is.na() will also come in handy. This example keeps the observations in df for which the variable x is not NA:filter(df, !is.na(x))    
```{r}
head(filter(hflights,hflights$Distance>=3000))
head(filter(hflights,UniqueCarrier %in% c("JetBlue", "Southwest", "Delta")))
head(filter(hflights,TaxiIn + TaxiOut>hflights$AirTime))
head(filter(hflights,DepTime<500|ArrTime>2200))
head(filter(hflights, DepDelay>0&ArrDelay<0))
head(filter(hflights, DepDelay>0&hflights$Cancelled==1))
```
arrange(), that reorders the rows according to single or multiple variables,    
```{r}
dtc <- filter(hflights, Cancelled == 1, !is.na(DepDelay))
head(arrange(dtc,DepDelay))
head(arrange(dtc,CancellationCode))
head(arrange(dtc,UniqueCarrier,DepDelay))
head(arrange(hflights, UniqueCarrier, desc(DepDelay)))
head(arrange(hflights, DepDelay + ArrDelay))
```
mutate(), used to add columns from existing data,   
mutate(df, z = x + y)   
```{r}
g1<-mutate(hflights,ActualGroundTime=ActualElapsedTime - AirTime)
g2<-mutate(g1,GroundTime=TaxiIn + TaxiOut)
g3<-mutate(g2,AverageSpeed=Distance / AirTime * 60)
```
你當然可以一次mutate很多個    
```{r}
m1 <- mutate(hflights, loss = ArrDelay - DepDelay,loss_ratio=loss/DepDelay)
m2<-mutate(hflights,TotalTaxi=TaxiIn+TaxiOut,ActualGroundTime=ActualElapsedTime- AirTime,Diff=TotalTaxi-ActualGroundTime)
```
summarise(), which reduces each group to a single row by calculating aggregate measures.    
```{r}
summarise(hflights, min_dist = min(Distance), max_dist = max(Distance))
summarise(filter(hflights, Diverted == 1), max_div = max(Distance))
```
Aggregate functions   
min(x) - minimum value of vector x.   
max(x) - maximum value of vector x.   
mean(x) - mean value of vector x.   
median(x) - median value of vector x.   
quantile(x, p) - pth quantile of vector x.    
sd(x) - standard deviation of vector x.   
var(x) - variance of vector x.    
IQR(x) - Inter Quartile Range (IQR) of vector x.    
diff(range(x)) - total range of vector x.   
```{r}
# Remove rows that have NA ArrDelay: temp1
temp1 <- filter(hflights, !is.na(ArrDelay))

# Generate summary about ArrDelay column of temp1
summarise(temp1, 
          earliest = min(ArrDelay), 
          average = mean(ArrDelay), 
          latest = max(ArrDelay), 
          sd = sd(ArrDelay))

# Keep rows that have no NA TaxiIn and no NA TaxiOut: temp2
temp2 <- filter(hflights, !is.na(TaxiIn), !is.na(TaxiOut))

# Print the maximum taxiing difference of temp2 with summarise()
summarise(temp2, max_taxi_diff = max(abs(TaxiIn - TaxiOut)))
```
first(x) - The first element of vector x.   
last(x) - The last element of vector x.   
nth(x, n) - The nth element of vector x.    
n() - The number of rows in the data.frame or group of observations that summarise() describes.   
n_distinct(x) - The number of unique values in vector x.    
```{r}
# Generate summarizing statistics for hflights
summarise(hflights, 
          n_obs = n(), 
          n_carrier = n_distinct(UniqueCarrier), 
          n_dest = n_distinct(Dest))

# All American Airline flights
aa <- filter(hflights, UniqueCarrier == "American")

# Generate summarizing statistics for aa 
summarise(aa, 
          n_flights = n(), 
          n_canc = sum(Cancelled == 1),
          avg_delay = mean(ArrDelay, na.rm = TRUE))
```
Pipe!!!%>%
```{r}
hflights %>%
mutate(diff=TaxiOut-TaxiIn) %>%
filter(!is.na(diff)) %>%
summarise(avg=mean(diff))

hflights%>%
mutate(RealTime=ActualElapsedTime + 100,mph=Distance/RealTime*60)%>%
filter(!is.na(mph),mph<70)%>%
summarise(n_less=n(),
n_dest=n_distinct(Dest),
min_dist=min(Distance),
max_dist=max(Distance)
)

hflights %>%
  mutate(RealTime = ActualElapsedTime + 100, mph = Distance / RealTime * 60)%>%
  filter(mph<105|Cancelled==1|Diverted==1)%>%
  summarise(
  n_non=n(),
  n_dest=n_distinct(Dest),
  min_dist=min(Distance),
  max_dist=max(Distance)
  )

hflights%>%
filter(!is.na(DepTime),!is.na(ArrTime),DepTime>ArrTime)%>%
summarise(num=n())
```
group_by()!!!
```{r}
hflights %>%
  group_by(UniqueCarrier) %>%
  summarise(p_canc = mean(Cancelled==1)*100,
            avg_delay = mean(ArrDelay,na.rm=1)) %>%
  arrange(avg_delay, p_canc)

hflights%>%
filter(!is.na(ArrDelay),ArrDelay>0)%>%
group_by(UniqueCarrier)%>%
summarise(avg=mean(ArrDelay))%>%
mutate(rank=rank(avg))%>%
arrange(rank)

# How many airplanes only flew to one destination?
hflights %>%
  group_by(TailNum) %>%
  summarise(ndest = n_distinct(Dest)) %>%
  filter(ndest == 1) %>%
  summarise(nplanes = n())

# Find the most visited destination for each carrier
hflights %>% 
  group_by(UniqueCarrier, Dest) %>%
  summarise(n = n()) %>%
  mutate(rank = rank(desc(n))) %>%
  filter(rank == 1)
```

